---
title: "Determine most common causes of failed accessions" 
output:   
  html_notebook: 
    toc: true
    toc_float: true
---

C. Savonen 7-1-19

Background of issues: #1358 and #642 of refine.bio 

### Question: What are the primary cause(s) of failure experiments failing to add to the API?

- "previous_human.txt" is the full list of human accessions that have been tried to add to the API. 
- "previous_rat.txt" is the full list of rat accessions that have been tried to add to the API. 

### Objectives: 

1) Obtain the list of full experiments in the API to date. 

2) Cross reference (by species) the full API experiment list with the lists of all accessions that have been tried: "previous_human.txt" and "previous_rat". Accessions that are in "previous" lists but NOT in the API are our list of "failed to add to the API"

3) Look for differences between "failed to add to API" and "successfully added to API". 
A not at all comprehensive list of potential suspects: 
* non-supported platforms
* non-supported technologies
* non-supported library strategies
* non-supported libraries

## Set up 

```{r Set Up}
`%>%` <- dplyr::`%>%`
```

## Obtain a full list of the experiments that are on the API

Set up the data frame we will store it in. 
Read the first page.

```{r}
# This will store the experiment info for all the api registered experiments
all_api_experiments <- data.frame()

# Read in that first page
seq_experiments_json <- jsonlite::fromJSON("https://api.refine.bio/experiments/")

# Start at page 1
seq_counter <- 1
```

Run the loop. 

```{r}
# While there's still another page, get more experiments!
while (!is.null(seq_experiments_json$`next`)) {
  # Print message so we know the progress
  message(paste0("RNA-seq experiment page ", seq_counter, "..."))
  
  # Skip this for the first page since we already read it in
  if (seq_counter > 1) {
    seq_experiments_json <- jsonlite::fromJSON(seq_experiments_json$`next`)
  }
  
  # Melt into a data.frame
  seq_api_experiments <- data.frame(cbind(seq_experiments_json$results$accession_code,
                                    seq_experiments_json$results$organisms,
                                    seq_experiments_json$results$technologies))
  
  # Add this set of experiments/organisms to the running data.frame
  all_api_experiments <- rbind.data.frame(all_api_experiments, seq_api_experiments)
  
  # Go on to the next page
  seq_counter <- seq_counter + 1 
}
```

Set up the data.frame with appropriate column names and save to a TSV.

```{r}
# Make the column names more sensible
colnames(all_api_experiments) <- c("acc", "organisms", "technology")

# Write these experiments and their info to a file
readr::write_tsv(file.path("api_registered_experiments.tsv"))
```
  
## Read in the "previous" lists 
  
```{r}
all_attempted_human_acc <- readr::read_delim("previous_human.txt", "\n", col_names = FALSE) %>%
  dplyr::pull("X1")
all_attempted_rat_acc  <- readr::read_delim("previous_rat.txt", "\n", col_names = FALSE) %>%
  dplyr::pull("X1")
```

## Cross reference the API list with the attempted list

Do this for human first: 

```{r}
# Extract just the human accessions
api_human_acc <- all_api_experiments %>% 
  dplyr::filter(grep(organisms, "HOMO_SAPIENS")) %>%
  dplyr::pull(acc)

# Failed human accessions
failed_human <- all_attempted_human_acc[!is.na(match(all_attempted_human_acc, api_human_acc))]
```

Now rat.

```{r}
# Extract just the rat accessions
api_rat_acc <- all_api_experiments %>% 
  dplyr::filter(grep(organisms, "RATTUS_NORVEGICUS")) %>%
  dplyr::pull(acc)

# Failed rat accessions
failed_rat <- all_attempted_rat_acc[!is.na(match(all_attempted_rat_acc, api_rat_acc))]
```

## Check that these lists make sense

```{r}

```

### Print session info

```{r}
sessionInfo()
```
