---
title: "Determine most common causes of failed accessions" 
output:   
  html_notebook: 
    toc: true
    toc_float: true
---

C. Savonen 7-1-19

Background of issues: #1358 and #642 of refine.bio 

### Question: What are the primary cause(s) of failure experiments failing to add to the API?

- "previous_human.txt" is the full list of human accessions that have been tried to add to the API. 
- "previous_rat.txt" is the full list of rat accessions that have been tried to add to the API. 

### Objectives: 

1) Obtain the list of full experiments in the API to date. 

2) Cross reference (by species) the full API experiment list with the lists of all accessions that have been tried: "previous_human.txt" and "previous_rat". Accessions that are in "previous" lists but NOT in the API are our list of "failed to add to the API"

3) Look for differences between "failed to add to API" and "successfully added to API". 
A not at all comprehensive list of potential suspects: 
* non-supported platforms
* non-supported technologies
* non-supported library strategies
* non-supported libraries

## Set up 

```{r Set Up}
# magrittr pipe
`%>%` <- dplyr::`%>%`

# Make a folder for the output here
if (!dir.exists("data")) {
  dir.create("data")
}
```

We will use this function when setting up the experiment data.

```{r}
# Make a function that collapses variable info into one row per accession
collapse_experiment <- function(variable) {
  # Collapses info by the experiment level so we can store one row per accession
  # in a data.frame
  #
  # Args:
  #   variable: A list formatted variable from the json read that has one list 
  #   per accession
  # Returns:
  #   An unlisted variable that has collapsed each experiment's info into one 
  #   string with comma separated values
  collapsed_info <- lapply(variable, function(exp_info) {
    paste0(exp_info, collapse = ",")
  })
  unlist(collapsed_info)
}
```

## Obtain a full list of the experiments that are on the API

Set up the data frame we will store it in. 
Read the first page.
If we already have the file though, we will skip this.

```{r}
# Only run this part if we haven't ran it and created the file before
if (file.exists(file.path("data", "api_registered_experiments.tsv"))) {
  all_api_experiments <- readr::read_tsv(file.path("data",
                                                   "api_registered_experiments.tsv"))
} else {
  # This will store the experiment info for all the api registered experiments
  all_api_experiments <- data.frame()

  # Read in that first page
  seq_experiments_json <- jsonlite::fromJSON("https://api.refine.bio/experiments/")

  # Start at page 1
  seq_counter <- 1
}
```

Run the loop. Unless we already have the file.

```{r}
# Only run this part if we haven't ran it before
if (!file.exists(file.path("data", "api_registered_experiments.tsv"))) {
  # While there's still another page, get more experiments!
  while (!is.null(seq_experiments_json$`next`)) {
    # Print message so we know the progress
    message(paste0("RNA-seq experiment page ", seq_counter, "..."))
  
    # Skip this for the first page since we already read it in
    if (seq_counter > 1) {
      seq_experiments_json <- jsonlite::fromJSON(seq_experiments_json$`next`)
    }
  
    # Melt into a data.frame
    seq_api_experiments <- 
      data.frame(acc = unlist(seq_experiments_json$results$accession_code),
                organisms = collapse_experiment(seq_experiments_json$results$organisms),
                technology = collapse_experiment(seq_experiments_json$results$technologies),
                platforms = collapse_experiment(seq_experiments_json$results$platforms),
                samples = collapse_experiment(seq_experiments_json$results$processed_samples))
  
    # Add this set of experiments/organisms to the running data.frame
    all_api_experiments <- rbind.data.frame(all_api_experiments, seq_api_experiments)
  
    # Go on to the next page
    seq_counter <- seq_counter + 1 
  }
}
```

Set up the data.frame with appropriate column names and save to a TSV.

```{r}
# Write these experiments and their info to a file
readr::write_tsv(all_api_experiments, 
                 file.path("data", "api_registered_experiments.tsv"))
```
  
## Read in the "previous" lists
We want to filter out non-RNA-seq experiment accessions and only keep unique
accessions. 

Do this for the human list: 
  
```{r}
# Read in the "previous list"
all_attempted_human_acc <- readr::read_delim("previous_human.txt", "\n", col_names = FALSE) %>%
  unique() %>%
  dplyr::pull("X1")
```

How many unique rat accessions are there?

```{r}
length(all_attempted_human_acc)
```

Do this for the rat list: 

```{r}
# Read in the "previous list"
all_attempted_rat_acc  <- readr::read_delim("previous_rat.txt", "\n", col_names = FALSE) %>%
  unique() %>%
  dplyr::pull("X1")
```

How many unique rat accessions are there?

```{r}
length(all_attempted_human_acc)
```

## Extract only the species' RNA-seq accessions

We are sorting by the accession number rather than the technology because we 
don't want microarray datasets that happen to have RNA-seq as well to show up. 
We only care about data from SRA.
We will write the accessions to a TSV file. 

Obtain this list for human experiments:

```{r}
# Extract just human RNA-seq accessions
api_human_acc <- all_api_experiments %>% 
  dplyr::filter(grepl("HOMO_SAPIENS", organisms)) %>%
  dplyr::filter(grepl("RNA-SEQ", technology)) %>%
  dplyr::filter(!grepl("GSE", acc)) %>%
  readr::write_tsv(file.path("data", "human_failed_survey_accessions.tsv"))

# Extract just the accessions
api_human_acc <- api_human_acc %>%
  dplyr::pull(acc)

# How many human RNA-seq accessions?
length(unique(api_human_acc))
```

Now rat.

```{r}
# Extract just the rat RNA-seq accessions
api_rat_acc <- all_api_experiments %>% 
  dplyr::filter(organisms == "RATTUS_NORVEGICUS") %>%
  dplyr::filter(grepl("RNA-SEQ", technology)) %>%
  dplyr::filter(!grepl("GSE", acc)) 

# Extract just the accessions
api_rat_acc <- api_rat_acc %>%
  dplyr::pull(acc)

# How many rat RNA-seq accessions?
length(unique(api_rat_acc))
```

## Cross reference these lists with the all_attempted lists. 
And write these to a TSV file. 

For human:

```{r}
# Failed human accessions
failed_human <- setdiff(all_attempted_human_acc, api_human_acc)

# How long is this list?
length(failed_human)
```

For rat: 

```{r}
# Failed rat accessions
failed_rat <- setdiff(all_attempted_rat_acc, api_rat_acc) 

# How long is this list?
length(failed_rat)
```

## Write the failed accessions to a TSV file

```{r}
# Write these to a TSV file
readr::write_tsv(data.frame(failed_human),
                 file.path("data", "human_failed_survey_accessions.tsv"))

readr::write_tsv(data.frame(failed_rat),
                 file.path("data", "rat_failed_survey_accessions.tsv"))
```

### Print session info

```{r}
sessionInfo()
```
